---
title: ""
author: "Mishkin Khunger, Sanat Lal, Sharmin Kantharia, Vishal Pathak"
date: "11/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r dataload}
library(ggplot2)


cd = read.csv("D:\\Data Science\\Cardio Vascular\\Cardio-Vascular\\cardio.csv")

dim(cd)

cd$age<- (cd$age/365)

cd$age<- round(cd$age) 
```

```{r basicchecks}

#Structure
str(cd)

#Summary
summary(cd)

# Anomalies found in height, weight, systolic (ap_hi) and diastolic (ap_lo)
#Gender,cholestrol,gluc,smoke,alco,active,cardio are needed to be converted into categorical variable

```
```{r lib}
library(pastecs)
library(psych)
library(Hmisc)
```

```{r Outlierdetectionofage}
fivenum(cd$age)
psych::describe(cd$age)
stat.desc(cd$age)
hist(cd$age,
     main = "Histogram of Age",
     xlab = "Age in Years")
boxplot(cd$age,
        main = toupper("Boxplot of Age"),
        ylab = "Age in years",
        col = "blue")
outlier_values <- boxplot.stats(cd$age)$out  # outlier values.
boxplot(cd$age, main="Age", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

```
As we can see skewness in -0.3 which means age is approximately symmetric. As if skewness is in between scale of (-0.5 to 0.5)it will be considered approximately symmetric.


#### so in age as we can see the outliers are value 30 which are outside of 1.5*IQR . We are going to use capping methiod in order to treat outlier. For missing values that lie outside the 1.5*IQR limits, we could cap it by replacing those observations outside the lower limit with the value of 5th %ile and those that lie above the upper limit, with the value of 95th %ile.

```{r Outliertreatmentofage}
qnt <- quantile(cd$age, probs=c(.25, .75), na.rm = T)
caps <- quantile(cd$age, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(cd$age, na.rm = T)
cd$age[cd$age < (qnt[1] - H)] <- caps[1]
cd$age[cd$age > (qnt[2] + H)] <- caps[2]

outlier_values <- boxplot.stats(cd$age)$out  # outlier values.
boxplot(cd$age, main="Age", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

``` 
#### As we can see now the age column has no outliers.we are good to go.

```{r HeightEDA}
fivenum(cd$height)
psych::describe(cd$height)
stat.desc(cd$height)

hist(cd$height,
     main = "Histogram of Height",
     xlab = "Height in centimeter")

outlier_values <- boxplot.stats(cd$cd$height)$out  # outlier values.
boxplot(cd$height, main="Height", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

outliersper <- function(x){
  length(which(x >  mean(x) + 3 * sd(x) | x < mean(x) - 3 * sd(x))  ) / length(x)
}

print(outliersper(cd$height))

```
#### As we can see there are few of outliers in Height variable , which is around .004 percent of total data. since this is continous data so we are going treat with capping method. For missing values that lie outside the 1.5*IQR limits, we could cap it by replacing those observations outside the lower limit with the value of 5th %ile and those that lie above the upper limit, with the value of 95th %ile. Moreover the skewness is around -0.63 which means height column is left skewed.

```{r Heightoutlierimpute}

qnt <- quantile(cd$height, probs=c(.25, .75), na.rm = T)
caps <- quantile(cd$height, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(cd$height, na.rm = T)
cd$height[cd$height < (qnt[1] - H)] <- caps[1]
cd$height[cd$height > (qnt[2] + H)] <- caps[2]

outlier_values <- boxplot.stats(cd$height)$out  # outlier values.
boxplot(cd$height, main="Height", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

```
#### As we can see now there are no outliers in height column.Also  earlier there was skewness in the column and it was (-0.63) but after the outlier treatment it reduces to 0.09 . so height variable is not skewed.so we are good to go.


```{r WeightEDA}
fivenum(cd$weight)
psych::describe(cd$weight)
stat.desc(cd$weight)

hist(cd$weight,
     main = "Histogram of Weight",
     xlab = "Weight in kg")

outlier_values <- boxplot.stats(cd$weight)$out  # outlier values.
boxplot(cd$weight, main="Weight", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

outliersper <- function(x){
  length(which(x >  mean(x) + 3 * sd(x) | x < mean(x) - 3 * sd(x))  ) / length(x)
}

print(outliersper(cd$weight))

```
#### As we can see there are 10 % of data which are outlier . Also data is highly asymmetric as skew value is (1.01) which meamns age is highly right skewed.we will first treat outlier ,then we will check if skewness persists then we will apply transformation on the weight column.



```{r Weightoutlierimpute}

qnt <- quantile(cd$weight, probs=c(.25, .75), na.rm = T)
caps <- quantile(cd$weight, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(cd$weight, na.rm = T)
cd$weight[cd$weight < (qnt[1] - H)] <- caps[1]
cd$weight[cd$weight > (qnt[2] + H)] <- caps[2]

outlier_values <- boxplot.stats(cd$height)$out  # outlier values.
boxplot(cd$height, main="Height", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

psych::describe(cd$weight)

```
#### As we can see now there are no outliers and also skewness is 0.39 which means our data is approximately normal.we are good to go.

```{r ap_hi(EDA)}
fivenum(cd$ap_hi)
psych::describe(cd$ap_hi)
stat.desc(cd$ap_hi)

hist(cd$ap_hi,
     main = "Histogram of systolic Blood Pressure",
     xlab = "Ap_hi")

outlier_values <- boxplot.stats(cd$ap_hi)$out  # outlier values.
boxplot(cd$ap_hi, main="systolic Blood Pressure", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

outliersper <- function(x){
  length(which(x >  mean(x) + 3 * sd(x) | x < mean(x) - 3 * sd(x))  ) / length(x)
}

print(outliersper(cd$ap_hi))

```
#### As we can see there is a 8 units of difference in mean and Meadian.which is really high . But the outliers are only around .05 % . This means that the values of outliers are of very values. Aslo the skew and kurtosis (85.29,7579.32)values are really very high. which again point to the fact that some values are incorrect or garbage values . we are to going treat the outliers in the next section and then we will analyse our column again.

```{r SBPoutlierimpute}

qnt <- quantile(cd$ap_hi, probs=c(.25, .75), na.rm = T)
caps <- quantile(cd$ap_hi, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(cd$ap_hi, na.rm = T)
cd$ap_hi[cd$ap_hi < (qnt[1] - H)] <- caps[1]
cd$ap_hi[cd$ap_hi > (qnt[2] + H)] <- caps[2]

outlier_values <- boxplot.stats(cd$ap_hi)$out  # outlier values.
boxplot(cd$ap_hi, main="systolic Blood Pressure", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

psych::describe(cd$ap_hi)

```

#### yes we were absolutely right the outliers were of very high and they were garbage value.After treating outliers , we can see our sweet boxplot(without any outliers) .Moreover our skew value is now under approximately normal range which is 0.54.

```{r ap_low(EDA)}
fivenum(cd$ap_lo)
psych::describe(cd$ap_lo)
stat.desc(cd$ap_lo)

hist(cd$ap_lo,
     main = "Histogram of Diastolic Blood Pressure",
     xlab = "Ap_low")

outlier_values <- boxplot.stats(cd$ap_lo)$out  # outlier values.
boxplot(cd$ap_lo, main="Diastolic Blood Pressure", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

outliersper <- function(x){
  length(which(x >  mean(x) + 3 * sd(x) | x < mean(x) - 3 * sd(x))  ) / length(x)
}

print(outliersper(cd$ap_lo))

```

#### As we can see there is difference of 16 units in mean and median.From the histogram we can see that data is right skewed which is around (32.11). Also the kurtosisis really high. there are around 10% outliers so we are going to treat it.after that we are going to check again

```{r DBDoutliertreatment}

qnt <- quantile(cd$ap_lo, probs=c(.25, .75), na.rm = T)
caps <- quantile(cd$ap_lo, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(cd$ap_lo, na.rm = T)
cd$ap_lo[cd$ap_lo < (qnt[1] - H)] <- caps[1]
cd$ap_lo[cd$ap_lo > (qnt[2] + H)] <- caps[2]

outlier_values <- boxplot.stats(cd$ap_lo)$out  # outlier values.
boxplot(cd$ap_lo, main="Diastolic Blood Pressure", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)

psych::describe(cd$ap_lo)

```
#### As we can see now after treatment of outliers , skew has also been in expected range(0.48) . now our variable is approximately normally distributed.

#Missing Value check and treatment on continous variables
```{r Missingvalues}

sum(is.na(cd$age))
sum(is.na(cd$height))
sum(is.na(cd$weight))
sum(is.na(cd$ap_hi))
sum(is.na(cd$ap_lo))

```

```{r genderEDA}
cd[cd$gender==1,]$gender<-"F"
cd[cd$gender==2,]$gender<-"M"
cd$gender<-as.factor(cd$gender)
table(factor(cd$gender))
```

#### There are no missing variables in the data , which means that we are good to go for our bivariate analysis.



```{r Checkaphivsaplo}

temp_min = pmin(cd$ap_hi, cd$ap_lo)
cd$ap_hi = pmax(cd$ap_hi, cd$ap_lo)
cd$ap_lo = temp_min
```
#### since diastolic and systolic are connected and diastolic will always be greater than systolic . so for sanity we have implement few changes to swap values if sysytolic is less than diastolic


```{r calculateBMI}

cd$bmi <- cd$weight/(cd$height*cd$height)*10000


```

```{r fac, include=FALSE}
#Gender
cd$gender<-as.factor(cd$gender)
#Cholestrol
cd$cholesterol<-as.factor(cd$cholesterol)
#Glucose
cd$gluc<-as.factor(cd$gluc)
#Smoke
cd$smoke<- as.factor(cd$smoke)
#Alcohol
cd$alco<- as.factor(cd$alco)
#Active
cd$active<- as.factor(cd$active)
#Cardio
cd$cardio<- as.numeric(cd$cardio)
```

```{r subsettingforfactor}
library(GPArotation)
l1 <- data.frame(cd$age ,cd$height,cd$weight , cd$ap_hi,cd$ap_lo , cd$bmi)
corrm<- cor(l1)     
```

```{r screeplot}
scree(corrm, factors=TRUE, pc=TRUE, main="scree plot", hline=NULL, add=FALSE) ### SCREE PLOT
eigen(corrm)$values                                                     ### EIGEN VALUES

```


#### As we can see the scree plot there are only 3 values which have eigen value above 1 , this means that we are going to use factor analysis for PC3.

```{r cumvarvar}
library(plyr)
eigen_values <- mutate(data.frame(eigen(corrm)$values)
                       ,cum_sum_eigen=cumsum(eigen.corrm..values)
                       , pct_var=eigen.corrm..values/sum(eigen.corrm..values)
                       , cum_pct_var=cum_sum_eigen/sum(eigen.corrm..values))
```

```{r factor analysis}
library(psych)
FA<-fa(r=corrm, 3, rotate="varimax", SMC=FALSE, fm="minres")
print(FA)                                                  
FA_SORT<-fa.sort(FA)                                        
ls(FA_SORT)                                                 
FA_SORT$loadings
```

#### From factor analysis we have taken 3 factors in account as per factor loadings. and those are bmi , api_hi , height. So we are good to go for our model implementation.




```{r cardioEDA}
cd[cd$cardio==0,]$cardio<-"Negative"
cd[cd$cardio==1,]$cardio<-"Positive"
cd$cardio<-as.factor(cd$cardio)
table(factor(cd$cardio))
```



#### There are no missing variables in the data , which means that we are good to go for our bivariate analysis.



## Calculating BMI

```{r bmi, echo=FALSE}
cd$bmi<- cd$weight/(cd$height*cd$height)*10000 # Weight in kgs and height in cms

```
## Dropping the id column
```{r drop id, echo=FALSE}
cd$id<- NULL


```

```{r levels}
levels(cd$gender)
levels(cd$cholesterol)
levels(cd$gluc)
levels(cd$smoke)
levels(cd$alco)
levels(cd$active)

```

```{r dummy variables}
cd$cholnormal  <- ifelse(cd$cholesterol == 1, 1, 0)
cd$cholabovenorm <- ifelse(cd$cholesterol == 2, 1, 0)
sum(cd$cholabovenorm)
sum(cd$cholnormal)

cd$glucnorm <- ifelse(cd$gluc == 1, 1, 0)
cd$glucabovenorm <- ifelse(cd$gluc == 2, 1, 0)
sum(cd$glucnorm)
sum(cd$glucabovenorm)

```

#### so we have create 2 dummy variables for cholestrol and 2 dummy variaables for glucose category and in both the columns there are 3 levels

```{r visualization1}
library(ggplot2)

#age count , affected vs not affected
counts <- table(cd$cardio, cd$age)
barplot(counts, main="Age distribution",
  xlab="Age", ylab="Counts", col=c("darkblue","red"),
  legend = rownames(counts), beside=TRUE)


```


#### We see that most people who are suffering from cardio vascular diseases are of the age 60, followed by 56.
Majorly, people belonging to the age group 50+ are suffering from the disease.



```{r v2}


#plot(cd$age,cd$ap_hi, main="Scatterplot Example",
 #  xlab="Car Weight ", ylab="Miles Per Gallon ")

library(car)
scatterplot(cd$ap_hi ~ cd$age, data=cd,
   xlab="Age", ylab="Systolic Blood pressure",
   main="Relationship of age and systolic blood pressure",
   )


mytable <- table(cd$cholesterol )
lbls <- paste(names(mytable), "\n", mytable, sep="")
pie(mytable, labels = lbls,
   main="Pie Chart of Cholestrol levels")

```


#### There are three cholestrol levels-
1- Normal
2- Above Normal
3- Way above normal

The above pie chart shows that 3/4th of the population come under the normal cholestrol category.We see that most people who are suffering from cardio vascular diseases are of the age 60, followed by 56.
Majorly, people belonging to the age group 50+ are suffering from the disease.


```{r v3}
library("ggplot2")

    ggplot(cd, aes(x= cholesterol,  group=cd$gender)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5) +
    labs(y = "Percent", fill="cholestrol") +
    facet_grid(~gender) +
    scale_y_continuous(labels = scales::percent)

```

#### The above graph shows that there are comparitively higher chances of women to have a cardio vascular disease where 76.9% of the women have normal cholestrol level.


```{r v4}

library("ggplot2")
attach(cd)
plot(gender, cardio, main="Scatterplot Example",
   xlab="gender ", ylab="Cardio ", pch=19)

counts <- table(cd$cardio, cd$gender)
barplot(counts, main="Gender distribution",
  xlab="Gender", ylab="Counts", col=c("pink","red"),
  legend = rownames(counts), beside=TRUE)

```

#### The above graph suggests that there are higher number of cases for women, but there are higher number of negative cases for women and higher number of positive cases for men.


```{r glvscd, include=FALSE, echo=TRUE}
  
  library("ggplot2")
  attach(cd)
  
  counts <- table(cd$cardio, cd$gluc)
  barplot(counts, main="Glucose distribution",
    xlab="Glucose", ylab="Counts", col=c("blue","black"),
    legend = rownames(counts), beside=TRUE)

```


```{r chcd, include=FALSE,echo=TRUE}
  
  library("ggplot2")
  attach(cd)

  
  counts <- table(cd$cardio, cd$cholesterol)
  barplot(counts, main="Cholestrol vs Cardio",
    xlab="Cholesterol", ylab="Counts", col=c("black","red"),
    legend = rownames(counts), beside=TRUE)
```
  
```{r dividedata, include=FALSE}
require(caTools) #install caTools
set.seed(123)
sample = sample.split(cd,SplitRatio = 0.70)
train_cd =subset(cd,sample ==TRUE)
test_cd =subset(cd, sample==FALSE)
```

#### we are dividing the data into test and train.


```{r stat, echo = TRUE}

# Relationship with the target variable


ttestresage = t.test(train_cd$age~train_cd$cardio)
ttestresage
```
####Strong significance

#### Statistical Significance between Height & Cardio Vascular Disease risk 
```{r hvscd}
ttestresheight = t.test(train_cd$height~train_cd$cardio)
ttestresheight
```
#### Comparitively weak significance

#### Statistical Significance between Weight & Cardio Vascular Disease risk 
```{r wvscd}
ttestresweight = t.test(train_cd$weight~train_cd$cardio)
ttestresweight
```
#### Strong significance

```{r bvscd, include=FALSE}
#Statistical Significance between BMI & Cardio Vascular Disease risk 
ttestresbmi = t.test(train_cd$bmi~train_cd$cardio)
ttestresbmi
```
#### Strong significance

#### Statistical Significance between Systolic & Cardio Vascular Disease 
```{r svdcd, include=FALSE}
ttestressys = t.test(train_cd$ap_hi~train_cd$cardio)
ttestressys 
```
#### Strong significance

#### Statistical Significance between Diastolic & Cardio Vascular Disease 
```{r dvscd, include=FALSE}
ttestresdia = t.test(train_cd$ap_lo~train_cd$cardio)
ttestresdia 
```
#### Strong significance



#### Statistical Significance between Cholestrol & Cardio Vascular Disease
```{r cvdcd, include=FALSE}
chisqrescholesterol = chisq.test(train_cd$cholesterol,train_cd$cardio)
chisqrescholesterol 
```
#### Strong significance

#### Statistical Significance between Glucose & Cardio Vascular Disease
```{r gvscd, include=FALSE}
chisqresgluc = chisq.test(train_cd$gluc,train_cd$cardio)
chisqresgluc 
```
#### Strong significance

#### Statistical Significance between Glucose & Cardio Vascular Disease
```{r avscd, include=FALSE}
chisqresalco = chisq.test(train_cd$alco,train_cd$cardio)
chisqresalco 
```
#### Not statistically significant but missing by a small margin

#### Statistical Significance between Physical Activity & Cardio Vascular Disease risk 
```{r scvscd, include=FALSE}
chisqresact = chisq.test(train_cd$active,train_cd$cardio)
chisqresact
```
#### Strong significance

```{r random forest}
library(randomForest)
set.seed(123)
rf_cd <- randomForest(cardio~.-weight-height, data=train_cd,
                   ntree = 500,
                   mtry = 3,
                   importance = TRUE)
print(rf_cd)
attributes(rf_cd)

```
#### Taking Age,Gender, Blood Pressure (Systolic and Diastolic), Alcohol,Smoking,Glucose,Active and BMI as predictors we get an Out of Bag ( set of boostrap datasets which does not contain  certain records from the original dataset) error rate of 26.68%


```{r oober, }
oob.error.data <- data.frame(
  Trees=rep(1:nrow(rf_cd$err.rate), times=3),
  Type=rep(c("OOB", "Negative", "Positive"), each=nrow(rf_cd$err.rate)),
  Error=c(rf_cd$err.rate[,"OOB"], 
    rf_cd$err.rate[,"Negative"], 
    rf_cd$err.rate[,"Positive"]))

ggplot(data=oob.error.data, aes(x=Trees, y=Error)) +
  geom_line(aes(color=Type))
```
#### Here we can see that error rate is the most stable around 500 trees for both the results as well as OOB. Therefore, we will continue with the  default value.

```{r oob, include=FALSE, echo=TRUE}
oob.rf_cd <- vector(length=10)
for(i in 1:10) {
  temp.model <- randomForest(cardio ~ .-weight-height, data=train_cd, mtry=i, ntree=500)
  oob.rf_cd[i] <- temp.model$err.rate[nrow(temp.model$err.rate),1]
}
oob.rf_cd
```

#### Here we can see that, we get the best result for mtry= 3 in terms of Out of Bag error rate

```{r pred1}
library(caret)
p1 <- predict(rf_cd, train_cd)
confusionMatrix(p1, train_cd$cardio)
```
#### The training set gives an accuracy of 76.86% 

```{r pred2}
library(caret)
p2 <- predict(rf_cd, test_cd)
confusionMatrix(p2, test_cd$cardio)
```

#### The test set gives an accuracy of 73.58%
```{r varimpo}
hist(treesize(rf_cd),
     main = "No. of Nodes for the Trees",
     col = "green")

# Variable Importance
varImpPlot(rf_cd,
           sort = T,
           n.var = 10,
           main = "Top 10 - Variable Importance")
importance(rf_cd)
varUsed(rf_cd)

```
#### The variable importance plot shows that in terms of Mean Decrease Gini value we get Systolic(Ap_hi), Diastolic(Ap_lo),bm, age and cholesterol to be our top 5 predictors


