---
title: "DS Project Sharmin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r dataload}
library(ggplot2)

cd = read.csv("cardio.csv")
dim(cd)
cd$age<- (cd$age/365)
cd$age<- round(cd$age) 
```

```{r basicchecks}
#Structure
str(cd)
#Summary
summary(cd)

# Anomalies found in height, weight, systolic (ap_hi) and diastolic (ap_lo)
#Gender,cholestrol,gluc,smoke,alco,active,cardio are needed to be converted into categorical variable
```

```{r datawrangling, echo=FALSE}
attach(cd) # attach a set of object to the search path directly 
#Fixing the negative values and change them to their absolute values

cd<-abs(cd) #negative systolic and diastolic value fixed 

#Fixing systolic and diastolic

#Systolic 
#The scale in most sphygmomanometer ranges from 0 to 400 
#Explanation for the below calculation
# If the systolic/diastolic in the data is below 10 then it is possibly an error and would be 
# If the systolic/diastolic in the data lies between 10 and 40 (most likely a typing error), then the data will be multiplied by 10
# If the systolic/diastolic in the data lies between 400 and 10000 (Out of range and possible typing errors), then the data will be divided by 100
# If the systolic/diastolic in the data is greater than 10000 (Out of range and possible typing errors), then the data will be divided by 100

correction<-function(cr){ifelse(cr<10,NA,
                        ifelse(cr>=10 & cr<40,cr*10,        
                        ifelse(cr>400 & cr<4000,cr/10,
                        ifelse(cr>4000 & cr<10000,cr/100,       
                        ifelse(cr>=10000,cr/100,cr)))))}
cd$ap_hi<- correction(cd$ap_hi)
cd$ap_lo<- correction(cd$ap_lo)

# Systolic cannot be lower than Diastolic, therefore we will swap the values where such is the case in the dataset
temp_min = pmin(cd$ap_hi, cd$ap_lo)
cd$ap_hi = pmax(cd$ap_hi, cd$ap_lo)
cd$ap_lo = temp_min

# fixing height - assuming the height of the population in females is not less than 150 cm(approx. 4 feet 11 inches) and not less than 155 cm(5 ft 1 inches) in males and not above 191 cm (6 Ft 2 inches) in females and not above 6 ft 6 inches in male . Replacing the value below these parameters by their respective medians 
female<-subset(cd,gender==1,select=c(gender,height,weight))
male<-subset(cd,gender==2,select=c(gender,height,weight))

maleheightm<-median(male$height)
femaleheightm<-median(female$height)

cd$height<-ifelse(cd$gender==1 & cd$height<150,femaleheightm,ifelse(cd$gender==1 & cd$height>191,femaleheightm, ifelse(cd$gender==2 & cd$height<155,maleheightm,ifelse(cd$gender==2 & cd$height>200,maleheightm,cd$height) ))) #??????

#cd$height


#fixing weight - assuming the weight of the population in females is not less than 45 kgs and not less than 55kgs  in males and both are not greater than 120Kgs. Replacing the value below these parameters by their respective means 

maleweightm<-median(male$weight)
femaleweightm<-median(female$weight)

cd$weight<-ifelse(cd$gender==1 & cd$weight<45,femaleweightm, ifelse(cd$gender==1 & cd$weight>120,femaleweightm,ifelse(cd$gender==2 & cd$weight<55,maleweightm,ifelse(cd$gender==2 & cd$weight>120,maleweightm,cd$weight)))) #??????


# Calculating BMI using weight and height. Potential predictor 

cd$bmi<- cd$weight/(cd$height*cd$height)*10000 # Weight in kgs and height in cms
#cd$bmi

# Check for Null Values
sum(is.na(cd))

#The abnormally low values in the case of systolic and and diastolic are converted to NA. As they are very low in number (32) and contribute to only 0.04% of the data, they can be dropped. 
cd<-na.omit(cd) 

# dropping id column 
cd$id<- NULL


# Converting data types
#Gender
cd$gender<-as.factor(cd$gender)
#Cholestrol
cd$cholesterol<-as.factor(cd$cholesterol)
#Glucose
cd$gluc<-as.factor(cd$gluc)
#Smoke
cd$smoke<- as.factor(cd$smoke)
#Alcohol
cd$alco<- as.factor(cd$alco)
#Active
cd$active<- as.factor(cd$active)
#Cardio
cd$cardio<- as.factor(cd$cardio)


#write.csv(cd,"D:\\Data Science\\Cardio Vascular\\test.csv", row.names = TRUE)
summary(cd)
```

``` {r datasplit, echo=FALSE}
require(caTools) #install caTools
set.seed(123)
sample = sample.split(cd,SplitRatio = 0.70)
train_cd =subset(cd,sample ==TRUE)
test_cd =subset(cd, sample==FALSE)
```

# Decision Tree

### We use the packages 'rpart' and 'rpart.plot' inorder to compute and plot the decision trees. The rpart() uses our target feature (cardio) against other features. 

In the frist tree, we compare the target against all the other features in the dataset. Using printcp() and plotcp() we can see the most important features we want to work with. As we can see, our most important deciding feature is ap_hi (systolic blood pressure). 

Using cross-validation, we can see the most important variables using complexity parameters. The optimized tree shows us that only one deciding feature is the basis to deicide whether a person has a CVD or not. However, we can also see that the oroginal tree has 5 most important features: ap_hi, ap_lo, cholestrol, age and bmi.

We use the packages 'rattle', 'rpart.plot' and 'RColorBrewer', to create a more clean tree with the different probablilities of the possiblilities of Yes and No of attaining CVD.

To check for accuracy, use the predict() and use the accuracy formula.

The accuracy of this decision tree is 71.66%


```{r DT, echo=FALSE}
#install.packages("party")
#install.packages("rpart")
library("rpart")
library("rpart.plot")

# Creating the tree
dt <- rpart(cardio ~ ., method = "class", data = train_cd) 
rpart.plot(dt, extra = 106)

# Plotting the tree
plot(dt)
text(dt, pretty = 0)

# Required libraries
#install.packages("rattle")
library("rattle")
library("rpart.plot")
library("RColorBrewer")

fancyRpartPlot(dt)

# Cross Validation 
printcp(dt)
plotcp(dt)
summary(dt)

# Accuracy

predict_dt <-predict(dt, test_cd, type = 'class')

table_mat <- table(test_cd$cardio, predict_dt)
table_mat

accuracy_dt <- sum(diag(table_mat)) / sum(table_mat)
accuracy_dt
```



